Performance and Development Review

### 2017-08-25 papers/Anja/PaperBody.tex
Probably the final revision of the thesis.

\cite{benarous:guionnet:2007}, 

@@

\section{The model}\label{sec:model:anja}\setcounter{equation}{0}
We consider a stochastic volatility model
\beam\label{eq:1a:anja}
X_{it}=\sigma_{it}\,Z_{it}\,,\qquad i,t\in\bbz\,,
\eeam
where $(Z_{it})$ is an iid field independent of a strictly stationary
ergodic field $(\sigma_{it})$ of non-negative \rv s.
We write $Z$, $\sigma$, $X$ for generic elements of the $Z$-,
$\sigma$- and $X$-fields \st\ $\sigma$ and $Z$ are independent.
A special case appears when $\sigma>0$ is a constant: then $(X_{it})$ constitutes an iid field.

@@

\section{Case (1): $Z$ dominates the tail}\label{sec:case1}\setcounter{equation}{0}

@@

\section{Case (2): $\sigma$ dominates the tail}\label{Sec:case2}\setcounter{equation}{0}
In this section we assume the conditions of Case (2); see Section \ref{subsec:Case:2}. Our goal is to derive results analogous to Case (1): \regvar\ 
of $(X_{it})$, infinite variance limits for $S_{ij}$ and limit theory for the eigenvalues of the corresponding sample 
covariance matrices. It turns out that this case offers a wider spectrum of possible limit behaviors and that we have to further distinguish our assumptions about the distribution of $\eta$. So, in addition to \eqref{Eq:tail:eta} we assume that either
\begin{equation}\label{Eq:inf:mom:eta} \E[\ex^{\eta \alpha}]=\infty\end{equation}
or 
\begin{equation}\label{Eq:conv:eq:eta} \lim_{x \to \infty}\frac{\P(\eta_1+\eta_2>x)}{\P(\eta_1>x)}=c \in (0,\infty) \;\; \Leftrightarrow\;\; \lim_{x \to \infty}\frac{\P(\ex^{\eta_1}\cdot \ex^{\eta_2}>x)}{\P(\ex^{\eta_1}>x)}=c \in (0,\infty)\end{equation}
hold, where $\eta_1$ and $\eta_2$ are independent copies of $\eta$.
\bre Following  Cline \cite{cline:1986}, we call 
the distribution of a \rv\ $\eta$ {\it convolution equivalent} if  $\ex^\eta$ is regularly varying and
relation \eqref{Eq:conv:eq:eta} holds. The assumptions \eqref{Eq:inf:mom:eta} and \eqref{Eq:conv:eq:eta} are mutually exclusive, since the only possible finite limit $c$ in \eqref{Eq:conv:eq:eta} is given by $c=2\E[\ex^{\eta \alpha}]$; see Davis and Resnick \cite{davis:resnick:1986}. There are, however, 
regularly varying distributions of $\ex^\eta$ which satisfy $\E[\ex^{\eta \alpha}]<\infty$ but not
\eqref{Eq:conv:eq:eta}. An example is given in Cline \cite{cline:1986}, p.\ 538;  see also Lemma~\ref{lem:product}(3) 
for a necessary and sufficient condition ensuring \eqref{Eq:conv:eq:eta}.
\ere
As we will see later, relations \eqref{Eq:inf:mom:eta} and \eqref{Eq:conv:eq:eta} cause
rather distinct limit behavior of the sample covariance matrix. In particular, \eqref{Eq:conv:eq:eta}
allows for non-vanishing off-diagonal elements of the normalized sample  covariance matrices, in contrast to Case (1). 
\par
For notational simplicity, define
\beao
\psi=\max_{k,l} \psi_{kl}\qquad \mbox{and}\qquad \Lambda=\{(k,l): \psi_{kl}=\psi\}\,.
\eeao
Recall that for convenience we assume that $\psi=1$; if the latter condition does not hold we can replace  
(without loss of generality) the \rv s $\eta_{kl}$ by $\psi \eta_{kl}$ and the coefficients $\psi_{kl}$ by $\psi_{kl}/\psi$.
For given $(i,j)$, we define 
\begin{equation} \label{eq:psi:ij:def}
\psi^{ij}= \max_{k,l}\, (\psi_{kl}+\psi_{k+i-j,l})\,.
\end{equation}
Notice that $1\le \psi^{ij}\le 2$. For $d\ge 1$, we write $\mathbf{i}=(i_1, \ldots, i_d), \mathbf{j}=(j_1, \ldots, j_d)$ for elements of $\mathbb{Z}^d$. For given  $\mathbf{i}$ and  $\mathbf{j}$ we also define
\beao
\psi^{\mathbf{i},\mathbf{j}}= \max_{1 \leq l \leq d} \psi^{i_l,j_l}.
\eeao

@@

\lefteqn{P\big((Y \Theta^{ij}_t)_{t=0,\ldots,n} \in B\big)}\\&=&\frac{\mu^{\bfX^{ij}}_n(\tilde{B})}{\mu^{\bfX^{ij}}_n(

@@

&=& \frac{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_1^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\nu_{\frac{\alpha}{\psi^{ij}}}(dy)\right]}{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\right]} \\
&=& \sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\frac{1}{| \Lambda_{i,j}^{(0)}|}\frac{\E\left[\I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})Y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]}{\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]},

@@

\section{Simulations and data example}\label{sec:simulation}\setcounter{equation}{0}

@@

\section{Some $\alpha$-stable limit theory}\setcounter{equation}{0}\label{App:A}
In this paper, we make frequently use of
Theorem 4.3 in Mikosch and Wintenberger \cite{mikosch:wintenberger:2016} which we quote for convenience:
\begin{theorem}\label{thm:mikwin}
Let $(\bfY_t)$ be an $\bbr^p$-valued strictly stationary \seq , $\bfS_n=\bfY_1+\cdots +\bfY_n$ and $(a_n)$ be \st\ $n\,\P(\|\bfY\|>a_n)\to 1$.
Also write for $\vep>0$, $\ov \bfY_t = \bfY_t \I(\|{\mathbf Y}_t\|\le \vep a_n)$, $\underline \bfY_t= \bfY_t-\ov  \bfY_t $ and
\beao
\ov \bfS_{l,n}= \sum_{t=1}^l \ov \bfY_t \qquad  \un \bfS_{l,n}= \sum_{t=1}^l \un \bfY_t\,.
\eeao
Assume the following conditions:
\begin{enumerate}
\item 
$(\bfY_t)$ is \regvary\ with index $\alpha\in (0,2)  \setminus\{1\}$ and spectral tail process $(\bfTh_j)$.
\item
A mixing condition holds: there exists an integer \seq\ $m_n\to\infty$ \st\ $k_n= [n/m_n]\to \infty$
and 
\beam\label{eq:chfa}
\E \ex^{i\bft'\underline \bfS_n/a_n} - \Big(\E \ex^{i\bft'\underline \bfS_{m_n,n}/a_n}\Big)^{k_n} \to 0\,,\qquad \nto\,,\qquad \bft\in\bbr^p\,.
\eeam
\item An anti-clustering condition holds:
\beam\label{eq:acl}
\lim_{l\to\infty} \limsup_{\nto} \P\big(\max_{t=l,\ldots,m_n} \|\bfY_t\|>\delta a_n\mid \|\bfY_0\|>\delta a_n\big)=0\,,\qquad \delta>0\,
\eeam
for the same sequence $(m_n)$ as in (2).
\item If $\alpha\in (1,2)$, in addition $\E[\mathbf{Y}]=\mathbf{0}$ and the vanishing small values condition holds: 
\beam\label{eq:vansm}
\lim_{\vep\downarrow 0}\limsup_{\nto} \P\big(a_n^{-1} \|\ov \bfS_n-\E [\ov \bfS_n]\|>\delta \big)=0\,,\qquad \delta>0\,
\eeam
and $\sum_{i=1}^\infty \E[\|\bfTh_i\|]<\infty$.
\end{enumerate}
Then $a_n^{-1} \bfS_n\std \xi_\alpha$ for an $\alpha$-stable $\bbr^p$-valued vector $\xi_\alpha$ with log characteristic function
\beam\label{eq:chfid}
\int_0^\infty \E\big[\ex^{i\,y\,\bft'\sum_{j=0}^\infty \bfTh_j}- \ex^{i\,y\,\bft'\sum_{j=1}^\infty \bfTh_j}-i\,y\,\bft'\I_{(1,2)}(\alpha)\big]\,
d(-y^{\alpha})\,,\qquad \bft\in\bbr^p\,.
\eeam
\end{theorem}\noindent 
\begin{remark}\label{rem:case:alpha:1}
If we additionally assume that $\mathbf{Y}$ is symmetric, which implies $\E[\ov \bfY]=\mathbf{0}$, then the statement 
of the theorem also holds for $\alpha=1$.
\end{remark}

@@

\begin{eqnarray*}  \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt) }{\P(Y_1^{a_{\max}}>t)}  &\geq& \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, \max_{j=1, \ldots, p}Y_j^{a_{\max}}\leq st) }{\P(Y_1^{a_{\max}}>t)} \\
&& + \sum_{j:a_j =a_{\max}} \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_j^{a_{\max}}>st) }{\P(Y_1^{a_{\max}}>t)}\\
&& -\frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_{j_1}^{a_{\max}}>st, Y_{j_2}^{a_{\max}}>st \mbox{ for some } j_1\neq j_2) }{\P(Y_1^{a_{\max}}>t)}, \;\; s>0.

@@

\begin{eqnarray*} && \lim_{t \to \infty}\frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}\\

@@

$$ \lim_{s \searrow 0}\lim_{t \to \infty} \frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}=\int_0^\infty \P\left(\left(\I(a_{ij}=a_{\max})z^{a_{\max}}\prod_{k \neq j}Y_k^{a_{ik}}\right)_{1 \leq i \leq n} \in B\right)\nu_{\alpha}(dz). $$

>>>>

\cite{arous:guionnet:2008}, 

@@

\section{The model}\label{sec:model:anja}
% \setcounter{equation}{0}
We consider a stochastic volatility model
\beam\label{eq:1a:anja}
X_{it}=\sigma_{it}\,Z_{it}\,,\qquad i,t\in\bbz\,,
\eeam
where $(Z_{it})$ is an iid field independent of a strictly stationary
ergodic field $(\sigma_{it})$ of non-negative \rv s.
We write $Z$, $\sigma$, $X$ for generic elements of the $Z$-,
$\sigma$- and $X$-fields \st\ $\sigma$ and $Z$ are independent.
A special case appears when $\sigma>0$ is a constant: then $(X_{it})$ constitutes an iid field.

@@

\section{Case (1): $Z$ dominates the tail}\label{sec:case1}
% \setcounter{equation}{0}

@@

\section{Case (2): $\sigma$ dominates the tail}\label{Sec:case2}
% \setcounter{equation}{0}
In this section we assume the conditions of Case (2); see Section \ref{subsec:Case:2}. Our goal is to derive results analogous to Case (1): \regvar\ 
of $(X_{it})$, infinite variance limits for $S_{ij}$ and limit theory for the eigenvalues of the corresponding sample 
covariance matrices. It turns out that this case offers a wider spectrum of possible limit behaviors and that we have to further distinguish our assumptions about the distribution of $\eta$. So, in addition to \eqref{Eq:tail:eta} we assume that either
\begin{equation}\label{Eq:inf:mom:eta} \E[\ex^{\eta \alpha}]=\infty\end{equation}
or 
\begin{equation}\label{Eq:conv:eq:eta} \lim_{x \to \infty}\frac{\P(\eta_1+\eta_2>x)}{\P(\eta_1>x)}=c \in (0,\infty) \;\; \Leftrightarrow\;\; \lim_{x \to \infty}\frac{\P(\ex^{\eta_1}\cdot \ex^{\eta_2}>x)}{\P(\ex^{\eta_1}>x)}=c \in (0,\infty)\end{equation}
hold, where $\eta_1$ and $\eta_2$ are independent copies of $\eta$.
\bre Following  Cline \cite{cline:1986}, we call 
the distribution of a \rv\ $\eta$ {\it convolution equivalent} if  $\ex^\eta$ is regularly varying and
relation \eqref{Eq:conv:eq:eta} holds. The assumptions \eqref{Eq:inf:mom:eta} and \eqref{Eq:conv:eq:eta} are mutually exclusive, since the only possible finite limit $c$ in \eqref{Eq:conv:eq:eta} is given by $c=2\E[\ex^{\eta \alpha}]$; see Davis and Resnick \cite{davis:resnick:1986}. There are, however, 
regularly varying distributions of $\ex^\eta$ which satisfy $\E[\ex^{\eta \alpha}]<\infty$ but not
\eqref{Eq:conv:eq:eta}. An example is given in Cline \cite{cline:1986}, p.\ 538;  see also Lemma~\ref{lem:product}(3) 
for a necessary and sufficient condition ensuring \eqref{Eq:conv:eq:eta}.
\ere
As we will see later, relations \eqref{Eq:inf:mom:eta} and \eqref{Eq:conv:eq:eta} cause
rather distinct limit behavior of the sample covariance matrix. In particular, \eqref{Eq:conv:eq:eta}
allows for non-vanishing off-diagonal elements of the normalized sample  covariance matrices, in contrast to Case (1). 
\par
For notational simplicity, define
\beao
\psi=\max_{k,l} \psi_{kl}\qquad \mbox{and}\qquad \Lambda=\{(k,l): \psi_{kl}=\psi\}\,.
\eeao
Recall that for convenience we assume that $\psi=1$; if the latter condition does not hold we can replace  
(without loss of generality) the \rv s $\eta_{kl}$ by $\psi \eta_{kl}$ and the coefficients $\psi_{kl}$ by $\psi_{kl}/\psi$.
For given $(i,j)$, we define 
\begin{equation} \label{eq:psi:ij:def}
\psi^{ij}= \max_{k,l}\, (\psi_{kl}+\psi_{k+i-j,l})\,.
\end{equation}
Notice that $1\le \psi^{ij}\le 2$. For $d\ge 1$, we write $\mathbf{i}=(i_1, \ldots, i_d), \mathbf{j}=(j_1, \ldots, j_d)$ for elements of $\mathbb{Z}^d$. For given  $\mathbf{i}$ and  $\mathbf{j}$ we also define
\beao
\psi^{\mathbf{i},\mathbf{j}}= \max_{1 \leq l \leq d} \psi^{i_l,j_l}.
\eeao

@@

\lefteqn{\P\big((Y \Theta^{ij}_t)_{t=0,\ldots,n} \in B\big)}\\
\frac{\mu^{\bfX^{ij}}_n(\tilde{B})}{\mu^{\bfX^{ij}}_n(

@@

\frac{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_1^\infty
\I_B\left(y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\I_{\Lambda_{i,j}^{(t)}}(u,v)\right)_{t=0, \ldots, n}|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\nu_{\frac{\alpha}{\psi^{ij}}}(dy)\right]}{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\right]} \\
\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\frac{1}{| \Lambda_{i,j}^{(0)}|}\frac{\E\left[
\I_B\left(y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\I_{\Lambda_{i,j}^{(t)}}(u,v)\right)_{t=0, \ldots, n}|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]}{\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]},

@@

\section{Simulations and data example}\label{sec:simulation}
% \setcounter{equation}{0}

@@

\section{Some $\alpha$-stable limit theory}
\label{App:A}
% \setcounter{equation}{0}
In this paper, we make frequently use of
Theorem 4.3 in Mikosch and Wintenberger \cite{mikosch:wintenberger:2016} which we quote for convenience:
\begin{theorem}\label{thm:mikwin}
Let $(\bfY_t)$ be an $\bbr^p$-valued strictly stationary \seq , $\bfS_n=\bfY_1+\cdots +\bfY_n$ and $(a_n)$ be \st\ $n\,\P(\|\bfY\|>a_n)\to 1$.
Also write for $\vep>0$, $\ov \bfY_t = \bfY_t \I(\|{\mathbf Y}_t\|\le \vep a_n)$, $\underline \bfY_t= \bfY_t-\ov  \bfY_t $ and
\beao
\ov \bfS_{l,n}= \sum_{t=1}^l \ov \bfY_t \qquad  \un \bfS_{l,n}= \sum_{t=1}^l \un \bfY_t\,.
\eeao
Assume the following conditions:
\begin{enumerate}
\item 
$(\bfY_t)$ is \regvary\ with index $\alpha\in (0,2)  \setminus\{1\}$ and spectral tail process $(\bfTh_j)$.
\item
A mixing condition holds: there exists an integer \seq\ $m_n\to\infty$ \st\ $k_n= [n/m_n]\to \infty$
and 
\beam\label{eq:chfa}
\E \ex^{i\bft'\underline \bfS_n/a_n} - \Big(\E \ex^{i\bft'\underline \bfS_{m_n,n}/a_n}\Big)^{k_n} \to 0\,,\qquad \nto\,,\qquad \bft\in\bbr^p\,.
\eeam
\item An anti-clustering condition holds:
\beam\label{eq:acl}
\lim_{l\to\infty} \limsup_{\nto} \P\big(\max_{t=l,\ldots,m_n} \|\bfY_t\|>\delta a_n\mid \|\bfY_0\|>\delta a_n\big)=0\,,\qquad \delta>0\,
\eeam
for the same sequence $(m_n)$ as in (2).
\item If $\alpha\in (1,2)$, in addition $\E[\mathbf{Y}]=\mathbf{0}$ and the vanishing small values condition holds: 
\beam\label{eq:vansm}
\lim_{\vep\downarrow 0}\limsup_{\nto} \P\big(a_n^{-1} \|\ov \bfS_n-\E [\ov \bfS_n]\|>\delta \big)=0\,,\qquad \delta>0\,
\eeam
and $\sum_{i=1}^\infty \E[\|\bfTh_i\|]<\infty$.
\end{enumerate}
Then $a_n^{-1} \bfS_n\std \xi_\alpha$ for an $\alpha$-stable $\bbr^p$-valued vector $\xi_\alpha$ with log characteristic function
\beam\label{eq:chfid}
\int_0^\infty \E\big[\ex^{i\,y\,\bft'\sum_{j=0}^\infty \bfTh_j}- \ex^{i\,y\,\bft'\sum_{j=1}^\infty \bfTh_j}-i\,y\,\bft'\I_{(1,2)}(\alpha)\big]\,
d(-y^{\alpha})\,,\qquad \bft\in\bbr^p\,.
\eeam
\end{theorem}\noindent 
\begin{remark}\label{rem:case:alpha:1}
If we additionally assume that $\mathbf{Y}$ is symmetric, which implies $\E[\ov \bfY]=\mathbf{0}$, then the statement 
of the theorem also holds for $\alpha=1$.
\end{remark}

@@

  && \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt) }{\P(Y_1^{a_{\max}}>t)}  \\
  &\geq&
  \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, \max_{j=1, \ldots, p}Y_j^{a_{\max}}\leq st) }{\P(Y_1^{a_{\max}}>t)} \\
+ \sum_{j:a_j =a_{\max}} \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_j^{a_{\max}}>st) }{\P(Y_1^{a_{\max}}>t)}\\
-\frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_{j_1}^{a_{\max}}>st, Y_{j_2}^{a_{\max}}>st \mbox{ for some } j_1\neq j_2) }{\P(Y_1^{a_{\max}}>t)}, \;\; s>0.

@@

    \lim_{t \to \infty}\frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}\\

@@

\lim_{s \searrow 0}\lim_{t \to \infty} \frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)} \\
\int_0^\infty \P\left(\left(\I(a_{ij}=a_{\max})z^{a_{\max}}\prod_{k \neq j}Y_k^{a_{ik}}\right)_{1 \leq i \leq n} \in B\right)
\nu_{\alpha}(dz).   
